\begin{acknowledgements}
\addchaptertocentry{\acknowledgementname} % Add the acknowledgements to the table of contents
This PhD project is the outcome of the work of many people: the core team was composed by Erika Coppola, Francesca Raffaele, Rita Nogherotto and myself. Graziano Giuliani and Fabio di Sante also were instrumental to several parts of the project thanks to their technical insight. 

Whenever fitting, specific acknowledgements are given inside the text or in footnotes, but here I would especially like to thank all of my colleagues at the ICTP Earth System Physics group, which funded the project and hosted me during my PhD.\\
In particular, I thank Rita, Francesca and Erika, for sharing this adventure with me: thanks for being patient when I was stubborn, for being understanding when I made mistakes.
I can only imagine it's not always easy to work with a pedantic and introvert person like myself.\\
Thanks to Graziano and Ivan, without whom the mysteries of computing would often have remained just that: mysteries.
I thank the colleagues and friends of a time now gone, Bianca, Csaba, Stefano, Ramon, Irene: our paths have crossed, and I hope they will again.

But most of all I want to thank those who, during these three years, more than colleagues have been friends.
Giorgio, for putting up with me all this time; Fabio, for teaching by example (to play table football, of course); Costanza, for that one ice cream, and for more to come; James, for being the most down to cheese guy; Rita once again, for the moral support and the reality checks in harsh times; Marco, for the smile always ready.\\
Thank you, friends, for the most important thing: because you made going to work not a chore, but a pleasure. 
% Well... most of the time, anyway!

I have a short memory. I am certain I am forgetting someone, and if that's you, reader, I beg your pardon. If you think you deserve my gratitude, well, you probably do: thank you.

\newpage
\section*{Author's note}
This thesis has been compressed for uploading and is missing hyperlinks.
The original version, containing full-resolution images and hyperlinks, can be obtained from \url{http://users.ictp.it/~afantini/PhD_thesis/PhD_thesis.pdf} or by requesting it to \href{mailto:adr.fantini@gmail.com}{adr.fantini@gmail.com}.


\end{acknowledgements}